Parallelizing Linear Recurrent Neural Nets Over Sequence Length

### 2018-04-24 poster/main.tex
add heatmap

\begin{alertblock}{Abstract}
RNN training and inference generally takes time linear in the sequence length because
of non-linear sequential dependencies.
We show the training and inference of RNNs with only linear
sequential dependencies can be parallelized over the sequence length using the
parallel scan algorithm, leading to rapid training on long sequences even with
small minibatch size. We use this insight and a parallel linear recurrence CUDA
kernel to accelerate several state of the art RNN architectures by up to 9x and
to solve a synthetic sequence classification task with a one million timestep
dependency.
\end{alertblock}

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------
\begin{block}{Background}
Large minibatches are necessary for computational performance but create large memory
requirements and may damage model generalization ability.

@@

Given $x_t$, $\lambda_t$ can compute $h_t=\lambda_t h_{t-1} + x_t$ for $t=1\ldots T$ on $p$
processors in $O(T/p + \log(p))$ with the classic parallel scan algorithm.
\end{block}

@@

\begin{figure}
\includegraphics[width=1.0\linewidth]{cumsum.png}
\caption{Parallel cumulative sum example}
\end{figure}

@@

\begin{block}{Gated Impulse Linear Recurrence}
We designed a recurrent layer that can take maximium advantage of parallel linear recurrence??

@@

RNNs have a transition function $s_t = f(s_{t-1},x_t)$. $s_t$ serves dual roles as a
summary of the past as well as the output of the unit. Non-linear $f$ in units such
as vanilla RNN and LSTM prevents parallelization over sequence length.

Replacing the summary of the past $s_{t-1}$ with a linear surrogate $\tilde{s}_{t-1}$
allows the easy adaption of any existing RNN architecture for parallel computation.

The state of an LSTM consists of $(c_t, h_t)$. $c_t$ is already computed
by linear recurrence, so a linear surrogate LSTM must only compute a
linear $\tilde{h}_t$. We let $h = \text{GILR}(x)$ and call it a GILR-LSTM.

@@

\begin{block}{Accelerating Existing Models}
  We implemented several recent linear RNNs with the parallel linear
  recurrence kernel, as well as a variant with purely serial evaluation.
  We were able to obtain significant speedups with parallel evaluation
  in all the architectures at longer sequence lengths.

  Since the parallel kernel computes the exact same result as the
  serial kernel, we obtain this speedup with no tradeoff in terms of
  reduced accuracy or precision.

  [Can be cut if no space]
  We used two stacked RNN layers with 256 hidden
  units, keeping the GPU memory usage constant by fixing $bT = 65,536$
  for minibatch size $b$ and sequence length $T$. QRNN(2) refers to a
  QRNN with filter size 2. %\citep{bradbury2017quasi}

\begin{table}[t]
\begin{center}
\begin{tabular}{lrrrr}
\label{table:rnn-throughput}
\small{Seq.\ Len.} & \small{SRU} & \small{QRNN(2)}
  & \small{QRNN(10)} & \small{GILR-LSTM}\\ \midrule
16 & 0.28 & 0.38 & 0.78 & 0.61\\
256 & 0.84 & 0.86 & 0.99 & 0.91\\
4,096 & 1.38 & 1.18 & 1.05 & 0.98\\
65,536 & 9.21 & 6.68 & 2.05 & 1.41\\ \bottomrule
\end{tabular}
\end{center}
\vspace{2ex}
\caption{Parallel kernel speedup for a variety of LS-RNNs.}
\end{table}
\end{block}

%----------------------------------------------------------------------------------------
%	RESULTS
%----------------------------------------------------------------------------------------

>>>>

\begin{alertblock}{Abstract}
RNN training and inference generally takes time linear in the sequence length because
of non-linear sequential dependencies.
We show the training and inference of RNNs with only linear
sequential dependencies can be parallelized over the sequence length using the
parallel scan algorithm, leading to rapid training on long sequences even with
small minibatch size. We use this insight and a parallel linear recurrence CUDA
kernel to accelerate several state of the art RNN architectures by up to 9x and
to solve a synthetic sequence classification task with a one million time step
dependency.
\end{alertblock}

%----------------------------------------------------------------------------------------
%	INTRODUCTION
%----------------------------------------------------------------------------------------
\begin{block}{Background}
Large minibatches are necessary for computational performance but create large memory
requirements and damage model generalization ability.

\vspace{1ex}

@@

\vspace{1ex}

Given $x_t$, $\lambda_t$ can compute $h_t=\lambda_t h_{t-1} + x_t$ for $t=1\ldots T$ on $p$
processors in $O(T/p + \log(p))$ with the classic parallel scan algorithm. Backpropagation
of gradient can also be parallelized with the same algorithm.
\end{block}

\vspace{2ex}

@@

\begin{figure}
\includegraphics[width=1.0\linewidth]{cumsum.png}
\caption{Cumulative sum parallelized over 3 processors}
\end{figure}

@@

\begin{block}{Gated Impulse Linear Recurrence}
Given a fast algorithm for evaluating linear recurrences, we introduce a new
linear recurrent layer called \textbf{gated impulse linear recurrence (GILR)}

@@

RNNs have a transition function $s_t = f(s_{t-1},x_t)$. $s_t$ serves dual roles as a
summary of the past as well as the output of the unit. Non-linear $f$ in units such
as vanilla RNN and LSTM prevents parallelization over sequence length.
\vspace{1ex}

Replacing the summary of the past $s_{t-1}$ with a linear surrogate $\tilde{s}_{t-1}$
allows the easy adaption of any existing RNN architecture for parallel computation.
Several existing RNNs can be viewed as linear surrogate RNNs.
\vspace{1ex}

The state of an LSTM consists of $(c_t, h_t)$. $c_t$ is already computed
by linear recurrence, so a linear surrogate LSTM must only compute a
linear $\tilde{h}_t$. A \textbf{GILR-LSTM} uses $\tilde{h} = \text{GILR}(x)$
\end{block}

@@

\begin{block}{Accelerating Existing Linear Surrogate RNNs}
We compare the performance of several recent linear RNNs
implemented with the parallel linear recurrence kernel against
a baseline serial linear recurrence kernel.

\vspace{1ex}
\begin{table}[t]
\begin{center}
\begin{tabular}{lrrrr}
\label{table:rnn-throughput}
\small{Seq.\ Len.} & \small{SRU} & \small{QRNN(2)}
  & \small{QRNN(10)} & \small{GILR-LSTM}\\ \midrule
16 & 0.28 & 0.38 & 0.78 & 0.61\\
256 & 0.84 & 0.86 & 0.99 & 0.91\\
4,096 & 1.38 & 1.18 & 1.05 & 0.98\\
65,536 & 9.21 & 6.68 & 2.05 & 1.41\\ \bottomrule
\end{tabular}
\end{center}
\vspace{2ex}
\caption{Parallel kernel speedup for a variety of LS-RNNs.   All models use two stacked RNN layers with 256 hidden
  units, keeping the GPU memory usage constant by fixing $bT = 65,536$
  for minibatch size $b$ and sequence length $T$. QRNN($k$) refers to a
  QRNN with filter size $k$. %\citep{bradbury2017quasi}
}
\end{table}
\end{block}

%----------------------------------------------------------------------------------------
\vspace{-1ex}
\begin{figure}
\includegraphics[width=0.9\linewidth]{cudnn_heatmap_gilr.png}
\caption{Throughput (thousand steps/s) for same sized LSTM and GILR-LSTM}
\end{figure}
